---
title: "Data2002 Project 1"
date: "`r Sys.Date()`"
author: "541035662 530029720 540872536 550505479 530210568 550041014"
format:
  html:
    #embed-resources: true # Creates a single HTML file as output
    code-fold: true # Show code chunks directly in the rendered HTML
    code-tools: true # Includes a menu to download the code file
    # code-tools are particularly important if you use inline R to
    # improve the reproducibility of your report
table-of-contents: true # (Optional) Creates a table of contents
number-sections: true # (Optional) Puts numbers next to heading/subheadings
---

# Introduction

This report examines how paid employment and study habits relate to students’ university experience, using survey data from DATA2002/2902 students. Three research questions investigate these relationships, with limitations including self-reported data and sampling bias.

## Data Wrangling

All data processing for this report was conducted in R (R Core Team 2026) within the Rmarkdown environment (Xie, Allaire, and Grolemund 2018) for reproducibility. Data wrangling was performed using base R and the tidyverse，Visualisations were produced with ggplot2 (Wickham et al. 2019).

Variables were cleaned and recoded, with missing values handled case-by-case. Import the survey dataset, rename variables for clarity, remove "Other" responses from `assess`, and remove `work_hrs` values above 70 hours per week. equivalent to it exceeding a standard full-time work week

boxplots comparing distributions before and after cleaning.save as "survey_cleaned.csv".


```{r}
#| message: false
library(tidyverse)

raw <- read_csv("DATA2X02_2026_2_survey_clean.csv", show_col_types = FALSE)

df <- raw %>%
  rename(
    age          = `What is your age? (years)`,
    height       = `What is your approximate height? (cms)`,
    intl_dom     = `Are you an International or Domestic student?`,
    unit         = `Are you in DATA2002 or DATA2902?`,
    grade1001    = `What grade did you get in DATA1001/1901 (or equivalent unit)?`,
    gradehope    = `What grade do you hope to get in DATA2002/2902?`,
    assess       = `When it comes to uni assessments, what do you tend to do?`,
    commute      = `On average, how long does it take you to get to campus? (mins) [1 direction]`,
    pay_rent     = `Do you pay rent?`,
    rent_amt     = `On average, how much rent do you pay per week? (dollars)`,
    work_hrs     = `On average, how many hours a week do you spend in paid employment? (hours)`,
    sleep_hrs    = `On average, how many hours sleep do you get per night? (hours)`,
    music        = `What is your favourite type of music?`,
    aliens       = `Do you believe in the existence of aliens? (Extraterrestial life)`,
    water        = `On average, how much water do you drink in a day? (litres)`,
    n_lang       = `How many spoken languages can you speak fluently?`,
    multitask    = `Do you multi-task in lectures? (ie do other tasks concurrently)`
  ) %>%
  filter(assess != "Other") %>%
  mutate(assess = factor(assess, levels = c("Immediately", "Gradually", "Cram")),
         multitask = factor(multitask, levels = c("no", "yes"), labels = c("No", "Yes")))

threshold <- 70

df_clean <- df %>% filter(work_hrs <= threshold)

ggplot(bind_rows(
    df %>% mutate(stage = "Before cleaning"),
    df_clean %>% mutate(stage = "After cleaning")
  ),
  aes(x = stage, y = work_hrs, fill = stage)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Before cleaning" = "#BFDCEB", "After cleaning" = "#F6D6A8")) +
  labs(
    title = "Weekly employment hours before and after cleaning",
    x = NULL, y = "Weekly employment hours", fill = NULL
  ) +
  theme_minimal()

write_csv(df_clean, "survey_cleaned.csv")
```

# Results

## RQ1: Is there a difference in weekly paid employment hours between students who cram before assessments and those who do not?
Paid employment is a key external time commitment that may shape how students approach assessment preparation (Kember 2004). We compare mean weekly work hours between students who cram before deadlines and those who do not, using a Welch two-sample t-test.

### Exploratory data analysis
```{r}
rq1_data <- df_clean %>%
  select(assess, work_hrs) %>%
  mutate(is_cram = if_else(assess == "Cram", "Cram", "Non-Cram"))
```
```{r}
rq1_data %>%
  group_by(is_cram) %>%
  summarise(
    n = n(),
    mean = mean(work_hrs),
    median = median(work_hrs),
    SD = sd(work_hrs),
    Q1 = quantile(work_hrs, 0.25),
    Q3 = quantile(work_hrs, 0.75),
    IQR = IQR(work_hrs),
    min = min(work_hrs),
    max = max(work_hrs)
  ) %>%
  knitr::kable(align = rep("c", 9))
```

```{r}
rq1_data %>%
  ggplot(aes(x = is_cram, y = work_hrs)) +
  geom_boxplot(
    aes(fill = is_cram),
    width = 0.5,
    outlier.shape = NA,
    alpha = 0.85
  ) +
  geom_jitter(
    aes(colour = "Individual observations"),
    position = position_jitter(width = 0.15, seed = 123),
    alpha = 0.35,
    size = 1.5
  ) +
  stat_summary(
    aes(colour = "Group mean"),
    fun = mean,
    geom = "point",
    shape = 18,
    size = 4
  ) +
  stat_summary(
    fun = mean,
    geom = "text",
    aes(label = after_stat(round(y, 2))),
    vjust = -0.75,
    size = 3.5
  ) +
  scale_fill_manual(
    name = "Cram status",
    values = c("Non-Cram" = "#BFE3DB", "Cram" = "#F3B5A7")
  ) +
  scale_colour_manual(
    name = NULL,
    values = c("Individual observations" = "#6F8FA3", "Group mean" = "#D9827A")
  ) +
  labs(
    x = NULL,
    y = "Weekly paid employment hours",
    title = "Weekly work hours by cramming behaviour",
    caption = "Diamonds and labels indicate group means."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "plain", hjust = 0.5),
    legend.position = "top",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )
```

Figure 1: Distribution of weekly paid employment hours by cramming status.

1. **Hypotheses:**
Let $\mu_1$ and $\mu_2$ be the population mean weekly work hours for Cram and Non-Cram students respectively. $H_0: \mu_1 = \mu_2$ vs $H_1: \mu_1 \neq \mu_2$.

$H_0: \mu_1 = \mu_2$  
$H_1: \mu_1 \neq \mu_2$

2. **Assumptions:**
The Welch t-test assumes approximate normality within each group, assessed using Shapiro-Wilk tests and QQ plots. Both p-values are below 0.05 and the QQ plots indicate non-normality, so a $\log(x+1)$ transformation is applied to reduce right-skewness. 
```{r}
ggplot(rq1_data, aes(sample = work_hrs, colour = is_cram)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ is_cram) +
  labs(
    title = "QQ plot of weekly work hours by Cram status",
    x = "Theoretical quantiles", y = "Sample quantiles"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```
```{r}
cram_raw <- rq1_data %>% filter(is_cram == "Cram") %>% pull(work_hrs)
noncram_raw <- rq1_data %>% filter(is_cram == "Non-Cram") %>% pull(work_hrs)
shapiro_cram_raw <- shapiro.test(cram_raw)
shapiro_noncram_raw <- shapiro.test(noncram_raw)
 
tibble::tibble(
  Group = c("Cram", "Non-Cram"),
  Statistic = c(shapiro_cram_raw$statistic, shapiro_noncram_raw$statistic),
  p_value = c(shapiro_cram_raw$p.value, shapiro_noncram_raw$p.value)
) %>%
  mutate(
    Statistic = round(Statistic, 3),
    p_value = format.pval(p_value, digits = 3, eps = 0.001)
  ) %>%
  knitr::kable(align = c("l", "c", "c"), caption = "Shapiro-Wilk test on raw work hours")  
```

```{r}
cram_log <- log1p(cram_raw)
noncram_log <- log1p(noncram_raw)
shapiro_cram_log <- shapiro.test(cram_log)
shapiro_noncram_log <- shapiro.test(noncram_log)
 
tibble::tibble(
  Group = c("Cram", "Non-Cram"),
  Statistic = c(shapiro_cram_log$statistic, shapiro_noncram_log$statistic),
  p_value = c(shapiro_cram_log$p.value, shapiro_noncram_log$p.value)
) %>%
  mutate(
    Statistic = round(Statistic, 3),
    p_value = format.pval(p_value, digits = 3, eps = 0.001)
  ) %>%
  knitr::kable(align = c("l", "c", "c"), caption = "Shapiro-Wilk test on log-transformed work hours")

manual_skew <- function(x) {
  n <- length(x)
  m <- mean(x)
  (sum((x - m)^3) / n) / (sum((x - m)^2) / n)^(3/2)
}
```
Normality is still not fully satisfied after transformation, although skewness falls substantially. The remaining non-normality is partly attributable to the large number of zero work-hour responses. Given the large sample sizes, the Central Limit Theorem supports the t-test approximation despite this departure, which is acknowledged as a limitation.
Skewness is calculated as:
$$\text{skewness} = \frac{\frac{1}{n}\sum(x_i - \bar{x})^3}{\left(\frac{1}{n}\sum(x_i - \bar{x})^2\right)^{3/2}}$$
```{r}
skew_cram_raw <- manual_skew(cram_raw)
skew_noncram_raw <- manual_skew(noncram_raw)
skew_cram_log <- manual_skew(cram_log)
skew_noncram_log <- manual_skew(noncram_log)
tibble::tibble(
  Group = c("Cram", "Non-Cram"),
  `Raw skewness` = round(c(skew_cram_raw, skew_noncram_raw), 2),
  `Log skewness` = round(c(skew_cram_log, skew_noncram_log), 2)
) %>%
  knitr::kable(align = c("l", "c", "c"), caption = "Skewness before and after log transformation")
```

3. **Test statistic:**
```{r}
welch_test <- t.test(cram_log, noncram_log, var.equal = FALSE)
welch_test
```
Test statistic: $t = 3.60$, $df = 167.28$. P-value: $p = 0.0004$. Cohen's $d = 0.372$, a small-to-medium effect.

4. **P-value:**
```{r}
#| label: rq1-effect-size
pooled_sd <- sqrt(((length(cram_log) - 1) * var(cram_log) +
                    (length(noncram_log) - 1) * var(noncram_log)) /
                    (length(cram_log) + length(noncram_log) - 2))
cohend <- (mean(cram_log) - mean(noncram_log)) / pooled_sd
cohend  
#Cohen's d = 0.372, indicating a small-to-medium effect size.
```
Cohen's d = 0.372, indicating a small-to-medium effect size.

5. **Conclusion:**
Since $p < 0.05$, we reject $H_0$. Cram students report significantly higher weekly work hours (log scale) than Non-Cram students, providing evidence that paid employment is associated with cramming behaviour.  

6. **Limitations:**
Both variables are self-reported, risking recall or social desirability bias. The 70-hour cleaning threshold is analyst-chosen. Normality was not fully achieved even after the log transformation, and reliance on the Central Limit Theorem for validity — rather than genuine normality — means the result is asymptotic rather than exact. As this is observational, cross-sectional data, the result establishes association, not causation — employment may displace study time, or procrastination-prone students may be more likely to seek work. Unmeasured confounders (course load, financial need) cannot be ruled out.

## RQ2: Multitasking and cramming

Our client, a university learning-support team, wants to know whether
multitasking during lectures is associated with cramming for assessments,
to help distinguish a lecture-engagement problem from a broader
time-management one.

```{r}
#| label: rq2-setup
#| message: false
rq2_data <- read_csv("survey_cleaned.csv") %>%
  select(assess, multitask) %>%
  drop_na(assess, multitask) %>%
  mutate(
    assess    = factor(assess, levels = c("Gradually", "Immediately", "Cram")),
    is_cram   = factor(
      if_else(assess == "Cram", "Cram", "Not Cram"),
      levels = c("Not Cram", "Cram")
    ),
    multitask = factor(multitask, levels = c("No", "Yes"))
  )

cram_tab <- table(rq2_data$is_cram, rq2_data$multitask)
chi_test <- chisq.test(cram_tab, correct = FALSE)

chi_statistic <- unname(chi_test$statistic)
chi_df        <- unname(chi_test$parameter)
chi_p         <- chi_test$p.value
min_expected  <- min(chi_test$expected)
n_total       <- sum(cram_tab)
cramers_v     <- sqrt(chi_statistic / n_total)

rq2_props <- rq2_data %>%
  count(is_cram, multitask) %>%
  group_by(is_cram) %>%
  mutate(prop = n / sum(n)) %>%
  filter(multitask == "Yes")
pct_cram    <- 100 * rq2_props$prop[rq2_props$is_cram == "Cram"]
pct_notcram <- 100 * rq2_props$prop[rq2_props$is_cram == "Not Cram"]
```

```{r}
#| label: fig-rq2-eda
#| fig-cap: "Proportion of students who multitask during lectures, by cramming behaviour."
rq2_data %>%
  count(is_cram, multitask) %>%
  group_by(is_cram) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = is_cram, y = prop, fill = multitask)) +
  geom_col(position = "fill", width = 0.5, alpha = 0.85) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(
    name = "Multitasks?",
    values = c("No" = "#A8DADC", "Yes" = "#E76F51")
  ) +
  labs(x = "Assessment preparation style", y = "Proportion of students") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.minor = element_blank())
```

@fig-rq2-eda shows multitasking is more common among students who cram
(`r sprintf("%.1f", pct_cram)`%) than those who don't
(`r sprintf("%.1f", pct_notcram)`%), motivating a formal test of this
association.

- **Hypothesis:** $H_0$: multitasking is independent of cramming vs
  $H_1$: multitasking is not independent of cramming.
- **Assumptions:** Observations are independent (one response per student),
  and expected cell counts must be $\geq 5$ in the $2\times2$ table --
  confirmed below (minimum $\approx$ `r sprintf("%.1f", min_expected)`), so
  the chi-squared approximation is valid.

```{r}
#| label: rq2-assumptions
chi_test$expected
```

- **Test statistic:** $T = \sum (O_{ij}-E_{ij})^2/E_{ij}$. Under $H_0$,
  $T \sim \chi^2_{`r chi_df`}$.

```{r}
#| label: rq2-test
chi_test
```

- **Observed test statistic:** $t_0 =$ `r sprintf("%.2f", chi_statistic)`
  with `r chi_df` degree of freedom.
- **p-value:** $P(\chi^2_{`r chi_df`} \geq$ `r sprintf("%.2f", chi_statistic)`$)
  =$ `r signif(chi_p, 3)`.
- **Decision:** Reject $H_0$ in favour of $H_1$ ($p < 0.001$). Students who
  cram are considerably more likely to multitask (Cramér's
  $V =$ `r sprintf("%.2f", cramers_v)`, a small-to-moderate effect). As this
  is a self-selected, self-reported survey, "cram" and "multitask" may be
  interpreted inconsistently across respondents, and the association does
  not imply causation -- a third factor such as overall workload could
  plausibly drive both behaviours.
## RQ3: Do students who multitask in class spend a different time per week on work compared to students who do not?

Paid employment is an external time commitment. Students balancing more work with study may exhibit different lecture behaviours. Following our previous research, we investigate whether paid-employment hours also differ between students who do and do not multitask.

```{r}
#| label: rq3-setup
#| message: false
df_clean <- read_csv("survey_cleaned.csv") %>%
  mutate(
    work_hrs = as.numeric(work_hrs),
    multitask = factor(multitask, levels = c("No", "Yes"))
  ) %>%
  filter(is.na(work_hrs) | work_hrs <= 70)
```

```{r}
#| label: tbl-flipper
#| tbl-cap: "Summary statistics of weekly paid employment hours by multitasking status."

rq3_summary <- df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  group_by(multitask) %>%
  summarise(
    n = n(),
    mean = mean(work_hrs),
    median = median(work_hrs),
    SD = sd(work_hrs),
    Q1 = quantile(work_hrs, 0.25),
    Q3 = quantile(work_hrs, 0.75),
    IQR = IQR(work_hrs),
    min = min(work_hrs),
    max = max(work_hrs)
  )

knitr::kable(
  rq3_summary,
  align = rep("c", ncol(rq3_summary))
)
```
@tbl-flipper and @fig-work-multitask show that multitasking students averaged 8.02 paid-work hours per week, compared with 6.23 hours for non-multitasking students---a difference of 1.79 hours. Both distributions are visually right-skewed, with many students reporting few or no hours of employment and a smaller number reporting substantially higher working hours.
```{r}
#| label: fig-work-multitask
#| fig-cap: "Distribution of weekly paid employment hours by multitasking status."

df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  ggplot(aes(x = multitask, y = work_hrs)) +

  geom_boxplot(
    aes(fill = multitask),
    width = 0.5,
    outlier.shape = NA,
    alpha = 0.4
  ) +

  geom_jitter(
    aes(colour = "Individual observations"),
    position = position_jitter(
      width = 0.15,
      seed = 123
    ),
    alpha = 0.35,
    size = 1.5
  ) +

  stat_summary(
    aes(colour = "Group mean"),
    fun = mean,
    geom = "point",
    shape = 18,
    size = 4
  ) +
  
  stat_summary(
    fun = mean,
    geom = "text",
    aes(label = after_stat(round(y, 2))),
    vjust = -0.75,
    size = 3.5
  ) +

  scale_fill_manual(
    name = "Multitasking",
    values = c(
      "No" = "#A8DADC",
      "Yes" = "#F4A261"
    )
  ) +
  
  scale_colour_manual(
    name = NULL,
    values = c(
      "Individual observations" = "#457B9D",
      "Group mean" = "red"
    )
  ) +

  labs(
    x = "Multitasking during lectures",
    y = "Weekly paid employment (hours)",
    title = "Weekly paid employment hours by multitasking status",
    caption = "Diamonds and labels indicate group means."
  ) +

  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )
```

```{r}
#| label: rq3-permutation-test

rq3_data <- df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  mutate(multitask = factor(multitask, levels = c("No", "Yes")))

welch_stat <- function(hours, group) {
  hours_yes <- hours[group == "Yes"]
  hours_no  <- hours[group == "No"]

  (mean(hours_yes) - mean(hours_no)) /
    sqrt(var(hours_yes) / length(hours_yes) +
           var(hours_no) / length(hours_no))
}

mean_yes <- mean(rq3_data$work_hrs[rq3_data$multitask == "Yes"])
mean_no  <- mean(rq3_data$work_hrs[rq3_data$multitask == "No"])
mean_difference <- mean_yes - mean_no
t_observed <- welch_stat(rq3_data$work_hrs, rq3_data$multitask)

set.seed(2002)
B <- 10000

t_permuted <- replicate(
  B,
  welch_stat(
    rq3_data$work_hrs,
    sample(rq3_data$multitask, replace = FALSE)
  )
)

n_yes <- sum(rq3_data$multitask == "Yes")
n_no  <- sum(rq3_data$multitask == "No")
sd_yes <- sd(rq3_data$work_hrs[rq3_data$multitask == "Yes"])
sd_no  <- sd(rq3_data$work_hrs[rq3_data$multitask == "No"])
pooled_sd <- sqrt(
  ((n_yes - 1) * sd_yes^2 + (n_no - 1) * sd_no^2) /
    (n_yes + n_no - 2)
)
cohens_d <- mean_difference / pooled_sd
```

```{r}
#| label: fig-rq3-permutation
#| fig-cap: "Permutation null distribution of the Welch t-statistic. Dashed red lines indicate the observed cut-offs at plus and minus the absolute observed statistic."

tibble(t_permuted = t_permuted) %>%
  ggplot(aes(x = t_permuted)) +
  geom_histogram(
    bins = 45,
    fill = "#8ECAE6",
    colour = "white"
  ) +
  geom_vline(
    xintercept = c(-abs(t_observed), abs(t_observed)),
    colour = "#D62828",
    linewidth = 0.9,
    linetype = "dashed"
  ) +
  labs(
    x = "Permuted Welch t-statistic",
    y = "Number of permutations",
    title = "Permutation distribution under the null hypothesis"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```
The two-sided permutation p-value is the proportion of simulated statistics at
least as extreme as the observed statistic:

```{r}
#| label: rq3-p-value

extreme_count <- sum(abs(t_permuted) >= abs(t_observed))

# Plus-one correction for a Monte Carlo permutation p-value
p_permutation <- (extreme_count + 1) / (B + 1)

tibble(
  extreme_permutations = extreme_count,
  total_permutations = B,
  permutation_p_value = p_permutation
) %>%
  knitr::kable(
    format = "html",
    digits = c(0, 0, 4),
    col.names = c(
      "Extreme permutations",
      "Total permutations",
      "Permutation p-value"
    ),
    align = c("c", "c", "c"),
    table.attr = 'class="table table-bordered table-striped" style="width: auto; margin-left: auto; margin-right: auto;"'
  )
```
1. **Hypotheses:** Let $\mu_Y$ and $\mu_N$ be the population mean weekly
   paid-employment hours for students who do and do not multitask,
   respectively. $H_0:\mu_Y-\mu_N=0$ versus
   $H_A:\mu_Y-\mu_N\neq0$.

2. **Assumptions:** Student responses are treated as independent, and the
   multitasking labels are exchangeable under $H_0$. Normality is unnecessary,
   although a mean-based statistic remains sensitive to large values.
   
3. **Test statistic:**
   $T=(\bar X_Y-\bar X_N)/\sqrt{S_Y^2/n_Y+S_N^2/n_N}$. The observed difference
   is `r sprintf("%.2f", mean_difference)` hours per week and
   $t_0=`r sprintf("%.3f", t_observed)`$.

4. **Permutation procedure:** With `set.seed(2002)`, the Yes/No labels were 
   shuffled $B=10{,}000$ times while employment hours were fixed. The Welch 
   statistic was recalculated to form @fig-rq3-permutation.

5. **P-value:** For the two-sided test, permutations satisfying
   $|T_b^*|\geq|t_0|$ were counted. The plus-one estimate is
   $p_{\mathrm{perm}}=`r sprintf("%.4f", p_permutation)`$.

6. **Decision:** Since $p<0.05$, we reject $H_0$. There is evidence that mean
   employment hours differ by multitasking status. multitasking students
   worked `r sprintf("%.2f", mean_difference)` more hours per week on average
   in this sample. However, Cohen's $d=`r sprintf("%.2f", cohens_d)`$ indicates
   a small effect, so its practical importance is limited. Both variables are
   self-reported and confounding may remain. As this is an observational survey 
   of students, the result cannot establish causation or automatically 
   generalize to others.

# Conclusion

Overall, the three analyses show that greater employment commitments were associated with more cramming and lecture multitasking, although the effects were modest.

These findings suggest that students’ employment and study habits may provide useful indicators of academic engagement and help teaching staff understand students’ study behaviours.

However, as the survey is observational and self-reported, the findings show associations rather than causation, with sampling bias, self-reporting, and confounding factors limiting generalisability.


# AI Usage Statement

AI was used as a supplementary tool to improve chart formatting and visualisation, help us understand the data, and provide ideas for analysis and presentation.


# ISI Professional Ethics Statement

[...]


# References
