---
title: "Data2002 Project 1"
date: "`r Sys.Date()`"
author: "541035662 530029720 540872536 550505479 530210568 550041014"
format:
  html:
    embed-resources: true
    code-fold: true
    code-tools: true
table-of-contents: true
number-sections: true
---

# Introduction

This report examines relationships among paid employment, cramming, and multitasking to understand how students’ work commitments relate to study habits. Data come from a survey of DATA2002/2902 students. Because students were not randomly sampled from the broader university population, selection bias limits generalisability. Self-reported work hours and study behaviours may also be affected by recall and social-desirability biases.

# Data Wrangling

All analyses were conducted in R (R Core Team, 2026) using the tidyverse (Wickham et al., 2019). Questions asking students to report “average” work hours and categorise study behaviours may be interpreted inconsistently because no reference period or precise behavioural definitions are provided. Missing values were removed separately for variables required in each analysis. Responses selecting “Other” for assessment preparation were excluded because they could not be consistently classified. Weekly paid-employment hours above 70 were excluded as extreme responses, informed by Australian working-hour standards of 38 hours plus reasonable additional hours (Fair Work Act 2009 (Cth) s 62).

```{r}
#| label: fig-cleaning
#| fig-cap: "Weekly employment hours before and after cleaning"
#| message: false

library(tidyverse)
raw <- read_csv("DATA2X02_2026_2_survey_clean.csv", show_col_types = FALSE)
df <- raw %>%
  rename(
    age          = `What is your age? (years)`,
    height       = `What is your approximate height? (cms)`,
    intl_dom     = `Are you an International or Domestic student?`,
    unit         = `Are you in DATA2002 or DATA2902?`,
    grade1001    = `What grade did you get in DATA1001/1901 (or equivalent unit)?`,
    gradehope    = `What grade do you hope to get in DATA2002/2902?`,
    assess       = `When it comes to uni assessments, what do you tend to do?`,
    commute      = `On average, how long does it take you to get to campus? (mins) [1 direction]`,
    pay_rent     = `Do you pay rent?`,
    rent_amt     = `On average, how much rent do you pay per week? (dollars)`,
    work_hrs     = `On average, how many hours a week do you spend in paid employment? (hours)`,
    sleep_hrs    = `On average, how many hours sleep do you get per night? (hours)`,
    music        = `What is your favourite type of music?`,
    aliens       = `Do you believe in the existence of aliens? (Extraterrestial life)`,
    water        = `On average, how much water do you drink in a day? (litres)`,
    n_lang       = `How many spoken languages can you speak fluently?`,
    multitask    = `Do you multi-task in lectures? (ie do other tasks concurrently)`
  ) %>%
  filter(assess != "Other") %>%
  mutate(assess = factor(assess, levels = c("Immediately", "Gradually", "Cram")),
         multitask = factor(multitask, levels = c("no", "yes"), labels = c("No", "Yes")))
threshold <- 70
df_clean <- df %>% filter(work_hrs <= threshold)
ggplot(bind_rows(
    df %>% mutate(stage = "Before cleaning"),
    df_clean %>% mutate(stage = "After cleaning")
  ),
  aes(x = stage, y = work_hrs, fill = stage)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Before cleaning" = "#BFDCEB", "After cleaning" = "#F6D6A8")) +
  labs(
    title = "Weekly employment hours before and after cleaning",
    x = NULL, y = "Weekly employment hours", fill = NULL
  ) +
  theme_minimal()
write_csv(df_clean, "survey_cleaned.csv")
```

# Results

## RQ1: Work hours and cramming

**Research Question:** Is there a difference in weekly paid employment hours between students who cram before assessments and those who do not?

Paid employment may affect study habits (Kember, 1999). We compare work hours between cram and non-cram students using the Wilcoxon rank-sum test.

```{r}
#| label: rq1-setup
rq1_data <- df_clean %>%
  select(assess, work_hrs) %>%
  mutate(is_cram = if_else(assess == "Cram", "Yes", "No"))
```

```{r}
#| label: tbl-rq1-table
#| tbl-cap: "Summary statistics of weekly paid employment hours by cramming status."

rq1_data %>%
  group_by(is_cram) %>%
  summarise(
    n = n(),
    mean = mean(work_hrs),
    median = median(work_hrs),
    SD = sd(work_hrs),
    Q1 = quantile(work_hrs, 0.25),
    Q3 = quantile(work_hrs, 0.75),
    IQR = IQR(work_hrs),
    min = min(work_hrs),
    max = max(work_hrs)
  ) %>%
  knitr::kable(align = rep("c", 9))
```

Means exceed medians and the Non-Cram median is zero: both groups are right-skewed with heavy zero-inflation.

```{r}
#| label: fig-rq1-eda
#| fig-cap: "Distribution of weekly paid employment hours by cramming status."

rq1_data %>%
  ggplot(aes(x = is_cram, y = work_hrs)) +
  geom_boxplot(
    aes(fill = is_cram),
    width = 0.5,
    outlier.shape = NA,
    alpha = 0.85
  ) +
  geom_jitter(
    aes(colour = "Individual observations"),
    position = position_jitter(width = 0.15, seed = 123),
    alpha = 0.35,
    size = 1.5
  ) +
  stat_summary(
    aes(colour = "Group median"),
    fun = median,
    geom = "point",
    shape = 18,
    size = 4
  ) +
  stat_summary(
    fun = median,
    geom = "text",
    aes(label = after_stat(round(y, 2))),
    vjust = -0.75,
    size = 3.5
  ) +
  scale_fill_manual(
    name = "Cram status",
    values = c("No" = "#BFE3DB", "Yes" = "#F3B5A7")
  ) +
  scale_colour_manual(
    name = NULL,
    values = c("Individual observations" = "#6F8FA3", "Group median" = "#D9827A")
  ) +
  labs(
    x = NULL,
    y = "Weekly paid employment hours",
    title = "Weekly work hours by cramming behaviour",
    caption = "Diamonds and labels indicate group medians."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "plain", hjust = 0.5),
    legend.position = "top",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )
```

@fig-rq1-eda confirms strong right-skew and zero-inflation, so normality cannot be assumed.

**1. Hypotheses:**

$H_0$: The distributions of weekly paid-employment hours are the same for Cram and Non-Cram students.

$H_1$: The distributions of weekly paid-employment hours differ between Cram and Non-Cram students.

**2. Assumptions:**

Samples are independent and the outcome is at least ordinal. Both groups are strongly right-skewed with substantial ties at zero (Table 2), so a Wilcoxon rank-sum test is used rather than relying on normality. The large sample size supports the normal approximation with continuity correction.

```{r}
#| label: tbl-rq1-skewness
#| tbl-cap: "Skewness and proportion of zero responses by cramming status."

cram_raw <- rq1_data %>% filter(is_cram == "Yes") %>% pull(work_hrs)
noncram_raw <- rq1_data %>% filter(is_cram == "No") %>% pull(work_hrs)

manual_skew <- function(x) {
  n <- length(x)
  m <- mean(x)
  (sum((x - m)^3) / n) / (sum((x - m)^2) / n)^(3/2)
}

tibble::tibble(
  Group = c("Yes", "No"),
  Skewness = round(c(manual_skew(cram_raw), manual_skew(noncram_raw)), 2),
  `% zero` = round(c(mean(cram_raw == 0), mean(noncram_raw == 0)) * 100, 1)
) %>%
  knitr::kable(align = c("l", "c", "c"))
```

**3. Test statistic:**

```{r}
#| label: tbl-rq1-wilcox
#| tbl-cap: "Wilcoxon rank-sum test comparing weekly paid employment hours by cramming status."

wilcox_test <- wilcox.test(cram_raw, noncram_raw, exact = FALSE, correct = TRUE)

tibble::tibble(
  Statistic = round(unname(wilcox_test$statistic), 1),
  p_value = format.pval(wilcox_test$p.value, digits = 3, eps = 0.001)
) %>%
  knitr::kable(align = c("c", "c"))
```

R reports the Mann-Whitney statistic; observed $w = `r round(unname(wilcox_test$statistic), 1)`$ (@tbl-rq1-wilcox).

```{r}
#| label: rq1-effect-size

n1 <- length(cram_raw)
n2 <- length(noncram_raw)
r_rb <- 1 - (2 * unname(wilcox_test$statistic)) / (n1 * n2)
```

**4. P-value:**

The observed \(p=0.000458\), with a small rank-biserial effect (\(|r|=0.189\)).

**5. Decision:**

We reject \(H_0\) (\(p<0.001\)), providing evidence that weekly paid-employment hours differ between Cram and Non-Cram students, with Cram students tending to report higher hours. The effect was small (\(|r|=0.189\)). As the data are observational, this indicates association rather than causation.


## RQ2: Multitasking and cramming

**Research Question:** Is multitasking during lectures associated with cramming before assessments?

Our client, a university learning-support team, wants to distinguish a lecture-engagement problem from a broader time-management one. Divided attention has been linked to weaker academic outcomes (Junco & Cotten, 2012), so we test whether it also travels with cramming using a chi-squared test of independence.

**Exploratory data analysis:**

Rows are dropped only for missing `assess` or `multitask`; the `work_hrs` threshold is not applied here, as neither variable involves paid employment.

```{r}
#| label: rq2-setup
rq2_data <- df %>%
  select(assess, multitask) %>%
  drop_na(assess, multitask) %>%
  mutate(is_cram = factor(
    if_else(assess == "Cram", "Yes", "No"),
    levels = c("No", "Yes")
  ))

cram_tab <- table(rq2_data$is_cram, rq2_data$multitask)
chi_test <- chisq.test(cram_tab, correct = FALSE)

chi_statistic <- unname(chi_test$statistic)
chi_df        <- unname(chi_test$parameter)
chi_p         <- chi_test$p.value
min_expected  <- min(chi_test$expected)
n_total       <- sum(cram_tab)
cramers_v     <- sqrt(chi_statistic / n_total)

yes_counts   <- cram_tab[c("Yes", "No"), "Yes"]
group_totals <- rowSums(cram_tab)[c("Yes", "No")]
prop_test    <- prop.test(x = yes_counts, n = group_totals, correct = FALSE)
pct_cram     <- 100 * unname(prop_test$estimate[1])
pct_notcram  <- 100 * unname(prop_test$estimate[2])
diff_ci      <- 100 * prop_test$conf.int

by_assess <- rq2_data %>%
  count(assess, multitask) %>%
  group_by(assess) %>%
  mutate(prop = 100 * n / sum(n)) %>%
  filter(multitask == "Yes")
pct_imm  <- by_assess$prop[by_assess$assess == "Immediately"]
pct_grad <- by_assess$prop[by_assess$assess == "Gradually"]
```

```{r}
#| label: tbl-rq2-table
#| tbl-cap: "Cramming behaviour by multitasking status, with marginal totals."
cram_tab %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  tibble::rownames_to_column("Preparation style") %>%
  knitr::kable(align = "c",
               col.names = c("Preparation style", "Does not multitask",
                             "Multitasks", "Total"))
```

```{r}
#| label: fig-rq2-eda
#| fig-cap: "Proportion of students who multitask during lectures, by cramming behaviour."
rq2_data %>%
  count(is_cram, multitask) %>%
  group_by(is_cram) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = is_cram, y = prop, fill = multitask)) +
  geom_col(position = "fill", width = 0.5, alpha = 0.85) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(
    name = "Multitasking",
    labels = c("No" = "Does not multitask", "Yes" = "Multitasks"),
    values = c("No" = "#BFE3DB", "Yes" = "#F3B5A7")
  ) +
  labs(x = "Assessment preparation style", y = "Proportion of students") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.minor = element_blank())
```

@tbl-rq2-table and @fig-rq2-eda show multitasking is more common among students who cram (`r sprintf("%.1f", pct_cram)`%) than among those who do not (`r sprintf("%.1f", pct_notcram)`%). The two non-cramming categories multitask at similar rates to each other (Immediately `r sprintf("%.1f", pct_imm)`%, Gradually `r sprintf("%.1f", pct_grad)`%), justifying their collapse into one group.

**1. Hypotheses:** $H_0$: multitasking is independent of cramming among DATA2x02 students, versus $H_1$: the two are not independent.

**2. Assumptions:** Observations are independent and every expected cell count must exceed 5; the minimum here is 51.5, so the approximation is valid without continuity correction (otherwise Fisher's exact test would apply).

```{r}
#| label: tbl-rq2-expected
#| tbl-cap: "Expected cell counts under independence."
chi_test$expected %>%
  round(1) %>%
  as.data.frame.matrix() %>%
  tibble::rownames_to_column("Preparation style") %>%
  knitr::kable(align = "c",
               col.names = c("Preparation style", "Does not multitask", "Multitasks"))
```

**3. Test statistic:** $$T=\sum_{i,j}\frac{(O_{ij}-E_{ij})^2}{E_{ij}} \sim \chi^2_{`r chi_df`} \text{ under } H_0$$ The observed value is $t_0 = `r sprintf("%.2f", chi_statistic)`$.

```{r}
#| label: rq2-test
chi_test
```

**4. P-value:** $P(\chi^2_{`r chi_df`} \geq `r sprintf("%.2f", chi_statistic)`) = `r signif(chi_p, 3)`$. Cramér's $V = `r sprintf("%.2f", cramers_v)`$ ($n = `r n_total`$), and the 95% CI for the difference in multitasking rates is `r sprintf("%.1f", diff_ci[1])` to `r sprintf("%.1f", diff_ci[2])` percentage points.

**5. Decision:** Reject $H_0$ ($p < 0.001$): students who cram are considerably more likely to multitask. The effect is modest, so multitasking is one correlate of cramming rather than a defining feature. As a self-selected, self-reported survey, "cram" and "multitask" may be interpreted inconsistently, and the association does not imply causation — overall workload could drive both.

## RQ3: Multitasking and employment hours

**Research Question:** Do students who multitask in class spend different amounts of time in paid employment?

Paid employment is an external time commitment that may compete with lecture engagement. We test whether paid-employment hours differ between students who multitask and those who do not.

```{r}
#| label: rq3-setup
#| message: false
df_rq3 <- read_csv("survey_cleaned.csv") %>%
  mutate(
    work_hrs = as.numeric(work_hrs),
    multitask = factor(multitask, levels = c("No", "Yes"))
  ) %>%
  filter(work_hrs <= 70)
```

```{r}
#| label: tbl-rq3-summary
#| tbl-cap: "Summary statistics of weekly paid employment hours by multitasking status."

rq3_summary <- df_rq3 %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  group_by(multitask) %>%
  summarise(
    n = n(),
    mean = mean(work_hrs),
    median = median(work_hrs),
    SD = sd(work_hrs),
    Q1 = quantile(work_hrs, 0.25),
    Q3 = quantile(work_hrs, 0.75),
    IQR = IQR(work_hrs),
    min = min(work_hrs),
    max = max(work_hrs)
  )

knitr::kable(
  rq3_summary,
  align = rep("c", ncol(rq3_summary))
)

rq3_mean_no <- rq3_summary$mean[rq3_summary$multitask == "No"]
rq3_mean_yes <- rq3_summary$mean[rq3_summary$multitask == "Yes"]
rq3_mean_difference <- rq3_mean_yes - rq3_mean_no
```

@tbl-rq3-summary and @fig-rq3-eda show multitasking students averaged `r sprintf("%.2f", rq3_mean_yes)` hours per week versus `r sprintf("%.2f", rq3_mean_no)` for non-multitasking students, a difference of `r sprintf("%.2f", rq3_mean_difference)` hours.

```{r}
#| label: fig-rq3-eda
#| fig-cap: "Distribution of weekly paid employment hours by multitasking status."

df_rq3 %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  ggplot(aes(x = multitask, y = work_hrs)) +
  geom_boxplot(
    aes(fill = multitask),
    width = 0.5,
    outlier.shape = NA,
    alpha = 0.4
  ) +
  geom_jitter(
    aes(colour = "Individual observations"),
    position = position_jitter(
      width = 0.15,
      seed = 123
    ),
    alpha = 0.35,
    size = 1.5
  ) +
  stat_summary(
    aes(colour = "Group mean"),
    fun = mean,
    geom = "point",
    shape = 18,
    size = 4
  ) +
  stat_summary(
    fun = mean,
    geom = "text",
    aes(label = after_stat(round(y, 2))),
    vjust = -0.75,
    size = 3.5
  ) +
  scale_fill_manual(
    name = "Multitasking",
    values = c(
      "No" = "#A8DADC",
      "Yes" = "#F4A261"
    )
  ) +
  scale_colour_manual(
    name = NULL,
    values = c(
      "Individual observations" = "#457B9D",
      "Group mean" = "red"
    )
  ) +
  labs(
    x = "Multitasking during lectures",
    y = "Weekly paid employment (hours)",
    title = "Weekly paid employment hours by multitasking status",
    caption = "Diamonds and labels indicate group means."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )
```

```{r}
#| label: rq3-permutation-test

rq3_data <- df_rq3 %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  mutate(multitask = factor(multitask, levels = c("No", "Yes")))

welch_stat <- function(hours, group) {
  hours_yes <- hours[group == "Yes"]
  hours_no  <- hours[group == "No"]

  (mean(hours_yes) - mean(hours_no)) /
    sqrt(var(hours_yes) / length(hours_yes) +
           var(hours_no) / length(hours_no))
}

mean_yes <- mean(rq3_data$work_hrs[rq3_data$multitask == "Yes"])
mean_no  <- mean(rq3_data$work_hrs[rq3_data$multitask == "No"])
mean_difference <- mean_yes - mean_no
t_observed <- welch_stat(rq3_data$work_hrs, rq3_data$multitask)

set.seed(2002)
B <- 10000

t_permuted <- replicate(
  B,
  welch_stat(
    rq3_data$work_hrs,
    sample(rq3_data$multitask, replace = FALSE)
  )
)

n_yes <- sum(rq3_data$multitask == "Yes")
n_no  <- sum(rq3_data$multitask == "No")
sd_yes <- sd(rq3_data$work_hrs[rq3_data$multitask == "Yes"])
sd_no  <- sd(rq3_data$work_hrs[rq3_data$multitask == "No"])
pooled_sd <- sqrt(
  ((n_yes - 1) * sd_yes^2 + (n_no - 1) * sd_no^2) /
    (n_yes + n_no - 2)
)
cohens_d <- mean_difference / pooled_sd
```



```{r}
#| label: rq3-p-value

extreme_count <- sum(abs(t_permuted) >= abs(t_observed))

# Plus-one correction for a Monte Carlo permutation p-value
p_permutation <- (extreme_count + 1) / (B + 1)

tibble(
  extreme_permutations = extreme_count,
  total_permutations = B,
  permutation_p_value = p_permutation,
  cohens_d = cohens_d
) %>%
  knitr::kable(
    format = "html",
    digits = c(0, 0, 4, 2),
    col.names = c(
      "Extreme permutations",
      "Total permutations",
      "Permutation p-value",
      "Cohen's d"
    ),
    align = c("c", "c", "c", "c"),
    table.attr = 'class="table table-bordered table-striped" style="width: auto; margin-left: auto; margin-right: auto;"'
  )
```

**1. Hypotheses:** $\mu_Y$, $\mu_N$: population mean weekly paid-employment hours for multitasking vs non-multitasking students. $H_0: \mu_Y = \mu_N$ versus $H_1: \mu_Y \neq \mu_N$.

**2. Assumptions:** Responses are independent. The studentized permutation test using the Welch statistic reduces reliance on normality and is more robust to unequal variances, although extreme observations may still influence the result.

**3. Test statistic:** $T=(\bar X_Y-\bar X_N)/\sqrt{S_Y^2/n_Y+S_N^2/n_N}$; observed $t_0=`r sprintf("%.3f", t_observed)`$.

```{r}
#| label: fig-rq3-permutation
#| fig-cap: "Permutation null distribution of the Welch t-statistic. Dashed red lines indicate the observed cut-offs at plus and minus the absolute observed statistic."

tibble(t_permuted = t_permuted) %>%
  ggplot(aes(x = t_permuted)) +
  geom_histogram(
    bins = 45,
    fill = "#8ECAE6",
    colour = "white"
  ) +
  geom_vline(
    xintercept = c(-abs(t_observed), abs(t_observed)),
    colour = "#D62828",
    linewidth = 0.9,
    linetype = "dashed"
  ) +
  labs(
    x = "Permuted Welch t-statistic",
    y = "Number of permutations",
    title = "Permutation distribution under the null hypothesis"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```

**4. P-value:** With `set.seed(2002)`, Yes/No labels were shuffled $B=10{,}000$ times while hours were fixed; permutations with $|T_b^*|\geq|t_0|$ gave $p_{\mathrm{perm}}=`r sprintf("%.4f", p_permutation)`$ from @fig-rq3-permutation using the plus-one estimate. Cohen's $d=`r sprintf("%.2f", cohens_d)`$ indicates a small effect.

**5. Decision:** Since $p=`r sprintf("%.4f", p_permutation)`<0.05$ (Phipson & Smyth, 2010), we reject $H_0$. There is evidence that mean employment hours differ by multitasking status; multitasking students worked `r sprintf("%.2f", mean_difference)` more hours per week on average in this sample. However, the effect is small, so its practical importance is limited. Both variables are self-reported and confounding may remain. As this is an observational survey of students, the result cannot establish causation or automatically generalise to others.


# Conclusion

Overall, paid employment, cramming, and multitasking were associated. Cram students tended to report higher weekly work hours and were more likely to multitask, while multitasking students reported 1.79 more work hours per week on average. However, effect sizes were small to modest. As the data are observational and self-reported, these findings indicate association rather than causation, and unmeasured confounding and the 70-hour cleaning threshold may affect the results.

# AI Usage Statement

AI was used as a supplementary tool to improve chart formatting and visualisation, help us understand the data, and provide ideas for analysis and presentation.


# ISI Professional Ethics Statement

We conducted the analysis responsibly by maintaining data integrity, applying appropriate statistical methods, and transparently reporting limitations. We avoided causal claims beyond the evidence and protected participants’ privacy by using survey data only for the intended academic analysis.


# References

Fair Work Act 2009 (Cth) s 62 (Austl.).

Junco, R., & Cotten, S. R. (2012). No A 4 U: The relationship between multitasking and academic performance. *Computers & Education, 59*(2), 505–514. https://doi.org/10.1016/j.compedu.2011.12.023

Kember, D. (1999). Integrating part-time study with family, work and social obligations. *Studies in Higher Education, 24*(1), 109–124.

Phipson, B., & Smyth, G. K. (2010). Permutation p-values should never be zero: Calculating exact p-values when permutations are randomly drawn. *Statistical Applications in Genetics and Molecular Biology, 9*(1), Article 39. https://doi.org/10.2202/1544-6115.1585

R Core Team. (2026). *R: A language and environment for statistical computing*. R Foundation for Statistical Computing.

Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J., Robinson, D., Seidel, D. P., Spinu, V., Takahashi, K., Vaughan, D., Wilke, C., Woo, K., & Yutani, H. (2019). Welcome to the tidyverse. *Journal of Open Source Software, 4*(43), 1686.

Xie, Y., Allaire, J. J., & Grolemund, G. (2018). *R Markdown: The Definitive Guide*. Chapman and Hall/CRC.
