---
title: "Data2002 Project 1"
date: "`r Sys.Date()`"
author: "541035662 530029720 540872536 550505479 530210568"
format: 
  html: 
    embed-resources: true # Creates a single HTML file as output
    code-fold: true # Code folding; allows you to show/hide code chunks
    code-tools: true # Includes a menu to download the code file 
    # code-tools are particularly important if you use inline R to 
    # improve the reproducibility of your report
table-of-contents: true # (Optional) Creates a table of contents
number-sections: true # (Optional) Puts numbers next to heading/subheadings
---

# Introduction

[Motivation + client]

[Dataset description]

[Survey questions used]

[Potential biases and limitations of the survey]


## Data Wrangling

[Import data]

[Variable selection]

[Missing-value handling]

[Recode cram variable]

[Clean paid employment hours]

**Load, rename, and clean**

Import the survey dataset, rename variables for clarity, remove "Other" responses from `assess`, and remove `work_hrs` values above 70 hours per week. This threshold was chosen because more than 70 hours is equivalent to a 10-hour day with no days off, exceeding a standard full-time work week and leaving little time for sleep, classes, or commuting.

**Compare and save**

Boxplot `work_hrs` before and after cleaning, then save the cleaned dataset as `survey_cleaned.csv`.

```{r}
library(tidyverse)

raw <- read_csv("DATA2X02_2026_2_survey_clean.csv", show_col_types = FALSE)

df <- raw %>%
  rename(
    age          = `What is your age? (years)`,
    height       = `What is your approximate height? (cms)`,
    intl_dom     = `Are you an International or Domestic student?`,
    unit         = `Are you in DATA2002 or DATA2902?`,
    grade1001    = `What grade did you get in DATA1001/1901 (or equivalent unit)?`,
    gradehope    = `What grade do you hope to get in DATA2002/2902?`,
    assess       = `When it comes to uni assessments, what do you tend to do?`,
    commute      = `On average, how long does it take you to get to campus? (mins) [1 direction]`,
    pay_rent     = `Do you pay rent?`,
    rent_amt     = `On average, how much rent do you pay per week? (dollars)`,
    work_hrs     = `On average, how many hours a week do you spend in paid employment? (hours)`,
    sleep_hrs    = `On average, how many hours sleep do you get per night? (hours)`,
    music        = `What is your favourite type of music?`,
    aliens       = `Do you believe in the existence of aliens? (Extraterrestial life)`,
    water        = `On average, how much water do you drink in a day? (litres)`,
    n_lang       = `How many spoken languages can you speak fluently?`,
    multitask    = `Do you multi-task in lectures? (ie do other tasks concurrently)`
  ) %>%
  filter(assess != "Other") %>%
  mutate(assess = factor(assess, levels = c("Immediately", "Gradually", "Cram")),
         multitask = factor(multitask, levels = c("no", "yes"), labels = c("No", "Yes")))

threshold <- 70

df_clean <- df %>% filter(work_hrs <= threshold)

ggplot(bind_rows(
    df %>% mutate(stage = "Before cleaning"),
    df_clean %>% mutate(stage = "After cleaning")
  ),
  aes(x = stage, y = work_hrs, fill = stage)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Before cleaning" = "#8ecae6", "After cleaning" = "#ffb703")) +
  labs(
    title = "Weekly employment hours before and after cleaning",
    x = NULL, y = "Weekly employment hours", fill = NULL
  ) +
  theme_minimal()

write_csv(df_clean, "survey_cleaned.csv")
```

# Results

## RQ1

[Motivation]

[EDA figure + summary statistics]

[Hypotheses]

[Assumptions]

[Test statistic]

[Observed test statistic]

[p-value]

[Decision + contextual conclusion]

[Limitations]


## RQ2
### Step 1: Data import & cleaning (multitasking vs cramming)

```{r}
#| label: rq2-setup
#| message: false
library(tidyverse)
```

```{r}
#| label: rq2-clean
survey <- read_csv("survey_cleaned.csv")

# Keep only the variables needed for RQ2, and derive a binary
# "Cram vs Not Cram" version of `assess` so the test lines up with the RQ
# statement ("multitasking and cramming"), rather than the original 3-level
# assessment-style variable (Gradually / Immediately / Cram).
rq2_data <- survey %>%
  select(assess, multitask) %>%
  drop_na(assess, multitask) %>%
  mutate(
    assess    = factor(assess, levels = c("Gradually", "Immediately", "Cram")),
    is_cram   = if_else(assess == "Cram", "Cram", "Not Cram"),
    is_cram   = factor(is_cram, levels = c("Not Cram", "Cram")),
    multitask = factor(multitask, levels = c("No", "Yes"))
  )

glimpse(rq2_data)
```

We only select `assess` and `multitask` (not the whole cleaned dataset), and
only drop rows missing on these two columns — so this subsetting is specific
to RQ2 and doesn't discard rows that are unusable for other RQs. `assess` is
collapsed from 3 levels into a binary `is_cram` variable to match the RQ2
question directly (cram vs not cram).

### Step 2: Exploratory visualisation (multitasking vs cramming)

```{r}
#| label: rq2-eda
#| fig-cap: "Proportion of students who multitask, by cramming behaviour"
rq2_data %>%
  count(is_cram, multitask) %>%
  group_by(is_cram) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = is_cram, y = prop, fill = multitask)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Assessment preparation style",
    y = "Proportion of students",
    fill = "Multitasks?",
    title = "Multitasking is more common among students who cram"
  ) +
  theme_minimal(base_size = 13)
```

Supplementary (optional) — the same relationship broken down by the original
3-level `assess` variable, useful as a descriptive figure even though the
formal test uses the collapsed 2x2 version:

```{r}
#| label: rq2-eda-3level
survey %>%
  drop_na(assess, multitask) %>%
  count(assess, multitask) %>%
  group_by(assess) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = assess, y = prop, fill = multitask)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Assessment preparation style (3 levels)",
       y = "Proportion of students", fill = "Multitasks?") +
  theme_minimal(base_size = 13)
```

Both plots show the same pattern: the multitasking rate is visibly higher
among students who cram (~75%) than among those who don't (~52% "Gradually",
~46% "Immediately"). This motivates formally testing the association in the
next step.

### Step 3: Hypothesis test — Chi-squared test of independence

**H:**
$H_0$: multitasking behaviour is independent of cramming behaviour.
$H_1$: multitasking behaviour is *not* independent of cramming behaviour.

```{r}
#| label: rq2-table
cram_tab <- table(rq2_data$is_cram, rq2_data$multitask)
cram_tab
```

**A (assumptions):** The chi-squared test of independence requires (i)
independent observations (one row per student — reasonable here, no student
appears twice in the survey) and (ii) expected cell counts $\geq 5$ in every
cell of the contingency table.

```{r}
#| label: rq2-assumptions
chi_test <- chisq.test(cram_tab, correct = FALSE)
chi_test$expected
```

All four expected counts are well above 5 (minimum ≈ 51.5), so the
chi-squared approximation is valid here — there is no need to fall back on
Fisher's exact test or a Yates continuity correction.

**T & P (test statistic & p-value):**

```{r}
#| label: rq2-test
chi_test
```

```{r}
#| label: rq2-effect-size
# Effect size (Cramer's V), to accompany the p-value
n <- sum(cram_tab)
cramers_v <- sqrt(unname(chi_test$statistic) / n)
cramers_v
```

Observed: $\chi^2 = 21.35$, $df = 1$, $p \approx 3.8\times10^{-6}$,
Cramér's $V \approx 0.18$ (small-to-moderate association).

**C (conclusion with limitations):**
There is very strong evidence ($p < 0.001$) against the null hypothesis of
independence: students who cram are considerably more likely to multitask
(75.2%) than students who don't (51.8%). However:

1. This is an observational, self-selected survey sample — voluntary
   response bias means busier or more disorganised students may be
   over- or under-represented.
2. "Cram" and "multitask" are both self-reported and self-interpreted
   categories, which may not be applied consistently across respondents.
3. Association is not causation — a third factor (e.g. workload,
   procrastination tendency) could plausibly drive both behaviours.
4. Collapsing `assess` from 3 levels to 2 (Cram vs Not Cram) loses some
   information about the distinction between "Gradually" and
   "Immediately" preparers.


## RQ3: Paid employment and multitasking

**Research question:** Do students who multitask during lectures spend a
different average number of hours per week in paid employment compared with
students who do not multitask?

Paid employment represents an external time commitment for students. Students balancing more paid work with university study may exhibit different lecture behaviours. Having examined paid employment in relation to cramming and multitasking in relation to cramming, we investigate whether paid-employment hours also differ between students who do and do not multitask during lectures.

```{r}
#| label: tbl-flipper
#| tbl-cap: "Summary statistics of weekly paid employment hours by multitasking status."

rq3_summary <- df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  group_by(multitask) %>%
  summarise(
    n = n(),
    mean = mean(work_hrs),
    median = median(work_hrs),
    SD = sd(work_hrs),
    Q1 = quantile(work_hrs, 0.25),
    Q3 = quantile(work_hrs, 0.75),
    min = min(work_hrs),
    max = max(work_hrs)
  )

knitr::kable(
  rq3_summary,
  align = rep("c", ncol(rq3_summary))
)
```
The boxplot in @fig-work-multitask compares weekly paid-employment hours between students who reported multitasking during lectures and those who did not. Students who reported multitasking had a higher mean weekly paid-employment time (8.02 hours) than those who did not multitask (6.23 hours), giving an observed mean difference of 1.79 hours per week. Both groups show visually right-skewed distributions, with many students reporting few or no hours of paid employment and a smaller number reporting substantially higher working hours. This observed difference in group means is examined formally using a permutation test.
```{r}
#| label: fig-work-multitask
#| fig-cap: "Distribution of weekly paid employment hours by multitasking status."

df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  ggplot(aes(x = multitask, y = work_hrs)) +
  
  geom_boxplot(
    aes(fill = multitask),
    width = 0.5,
    outlier.shape = NA,
    alpha = 0.4
  ) +
  
  geom_jitter(
    aes(colour = "Individual observations"),
    position = position_jitter(
      width = 0.15,
      seed = 123
    ),
    alpha = 0.35,
    size = 1.5
  ) +
  
  stat_summary(
    aes(colour = "Group mean"),
    fun = mean,
    geom = "point",
    shape = 18,
    size = 4
  ) +
  
  stat_summary(
    fun = mean,
    geom = "text",
    aes(label = after_stat(round(y, 2))),
    vjust = -0.75,
    size = 3.5
  ) +
  
  scale_fill_manual(
    name = "Multitasking",
    values = c(
      "No" = "#A8DADC",
      "Yes" = "#F4A261"
    )
  ) +
  
  scale_colour_manual(
    name = NULL,
    values = c(
      "Individual observations" = "#457B9D",
      "Group mean" = "red"
    )
  ) +
  
  labs(
    x = "Multitasking during lectures",
    y = "Weekly paid employment (hours)",
    title = "Weekly paid employment hours by multitasking status",
    caption = "Diamonds and labels indicate group means."
  ) +
  
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )
```

```{r}
#| label: rq3-permutation-test
#| include: false

rq3_data <- df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  mutate(multitask = factor(multitask, levels = c("No", "Yes")))

welch_stat <- function(hours, group) {
  hours_yes <- hours[group == "Yes"]
  hours_no  <- hours[group == "No"]

  (mean(hours_yes) - mean(hours_no)) /
    sqrt(var(hours_yes) / length(hours_yes) +
           var(hours_no) / length(hours_no))
}

mean_yes <- mean(rq3_data$work_hrs[rq3_data$multitask == "Yes"])
mean_no  <- mean(rq3_data$work_hrs[rq3_data$multitask == "No"])
mean_difference <- mean_yes - mean_no
t_observed <- welch_stat(rq3_data$work_hrs, rq3_data$multitask)

set.seed(2002)
B <- 10000

t_permuted <- replicate(
  B,
  welch_stat(
    rq3_data$work_hrs,
    sample(rq3_data$multitask, replace = FALSE)
  )
)

n_yes <- sum(rq3_data$multitask == "Yes")
n_no  <- sum(rq3_data$multitask == "No")
sd_yes <- sd(rq3_data$work_hrs[rq3_data$multitask == "Yes"])
sd_no  <- sd(rq3_data$work_hrs[rq3_data$multitask == "No"])
pooled_sd <- sqrt(
  ((n_yes - 1) * sd_yes^2 + (n_no - 1) * sd_no^2) /
    (n_yes + n_no - 2)
)
cohens_d <- mean_difference / pooled_sd
```

```{r}
#| label: fig-rq3-permutation
#| fig-cap: "Permutation null distribution of the Welch t-statistic. Dashed red lines indicate the observed cut-offs at plus and minus the absolute observed statistic."

tibble(t_permuted = t_permuted) %>%
  ggplot(aes(x = t_permuted)) +
  geom_histogram(
    bins = 45,
    fill = "#8ECAE6",
    colour = "white"
  ) +
  geom_vline(
    xintercept = c(-abs(t_observed), abs(t_observed)),
    colour = "#D62828",
    linewidth = 0.9,
    linetype = "dashed"
  ) +
  labs(
    x = "Permuted Welch t-statistic",
    y = "Number of permutations",
    title = "Permutation distribution under the null hypothesis"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```

The two-sided permutation p-value is calculated by counting simulated
statistics at least as extreme as the observed statistic in either direction:

```{r}
#| label: rq3-p-value

extreme_count <- sum(abs(t_permuted) >= abs(t_observed))

# Plus-one correction for a Monte Carlo permutation p-value
p_permutation <- (extreme_count + 1) / (B + 1)

tibble(
  extreme_permutations = extreme_count,
  total_permutations = B,
  permutation_p_value = p_permutation
) %>%
  knitr::kable(
    format = "html",
    digits = c(0, 0, 4),
    col.names = c(
      "Extreme permutations",
      "Total permutations",
      "Permutation p-value"
    ),  
    align = c("c", "c", "c"),
    table.attr = 'class="table table-bordered table-striped" style="width: auto; margin-left: auto; margin-right: auto;"'
  )
```

We formally test the difference using a two-sided permutation test with the
Welch $t$-statistic as the resampling statistic.

1. **Hypotheses:** Let $\mu_Y$ and $\mu_N$ be the population mean weekly
   paid-employment hours for students who do and do not multitask,
   respectively. $H_0:\mu_Y-\mu_N=0$ versus
   $H_A:\mu_Y-\mu_N\neq0$.

2. **Assumptions:** Responses from different students are treated as
   independent. Under $H_0$, multitasking labels are assumed exchangeable. A
   normal population is not required for the permutation p-value, although the
   mean-based statistic remains sensitive to large values.

3. **Test statistic:**
   $T=(\bar X_Y-\bar X_N)/\sqrt{S_Y^2/n_Y+S_N^2/n_N}$. The observed difference
   is `r sprintf("%.2f", mean_difference)` hours per week and
   $t_0=`r sprintf("%.3f", t_observed)`$.

4. **Permutation procedure:** With `set.seed(2002)`, the Yes/No labels were
   randomly reassigned $B=10{,}000$ times while employment hours were held
   fixed. The Welch statistic was recalculated each time to form the null
   distribution in @fig-rq3-permutation.

5. **P-value:** For the two-sided test, permutations satisfying
   $|T_b^*|\geq|t_0|$ were counted. The plus-one estimate is
   $p_{\mathrm{perm}}=`r sprintf("%.4f", p_permutation)`$.

6. **Decision:** Since $p<0.05$, we reject $H_0$. There is evidence that mean
   paid-employment hours differ by multitasking status; multitasking students
   worked `r sprintf("%.2f", mean_difference)` more hours per week on average
   in this sample. However, Cohen's $d=`r sprintf("%.2f", cohens_d)`$ indicates
   a small effect, so its practical importance is limited. Both variables are
   self-reported, the 70-hour cleaning threshold is analyst-chosen, and
   confounding may remain. As this is an observational survey of DATA2x02
   students, the result cannot establish causation or automatically generalise
   to other students.

# Conclusion

[Integrate the three findings into one story]

[Implications for client]

[Overall limitations / no causal inference]


# AI Usage Statement

[...]


# ISI Professional Ethics Statement

[...]


# References