---
title: "Data2002 Project 1"
date: "`r Sys.Date()`"
author: "541035662 530029720 540872536 550505479 530210568 550041014"
format:
  html:
    #embed-resources: true # Creates a single HTML file as output
    code-fold: true # Show code chunks directly in the rendered HTML
    code-tools: true # Includes a menu to download the code file
    # code-tools are particularly important if you use inline R to
    # improve the reproducibility of your report
table-of-contents: true # (Optional) Creates a table of contents
number-sections: true # (Optional) Puts numbers next to heading/subheadings
---

# Introduction

[Motivation + client]

[Dataset description]

[Survey questions used]

[Potential biases and limitations of the survey]


## Data Wrangling

All data processing for this report was conducted in R (R Core Team 2026) within the Rmarkdown environment (Xie, Allaire, and Grolemund 2018) for reproducibility. Data wrangling was performed using base R and the tidyverse，Visualisations were produced with ggplot2 (Wickham et al. 2019).

Variables were cleaned and recoded, with missing values handled case-by-case. Import the survey dataset, rename variables for clarity, remove "Other" responses from `assess`, and remove `work_hrs` values above 70 hours per week. equivalent to it exceeding a standard full-time work week

boxplots comparing distributions before and after cleaning.save as "survey_cleaned.csv".


```{r}
#| message: false
library(tidyverse)

raw <- read_csv("DATA2X02_2026_2_survey_clean.csv", show_col_types = FALSE)

df <- raw %>%
  rename(
    age          = `What is your age? (years)`,
    height       = `What is your approximate height? (cms)`,
    intl_dom     = `Are you an International or Domestic student?`,
    unit         = `Are you in DATA2002 or DATA2902?`,
    grade1001    = `What grade did you get in DATA1001/1901 (or equivalent unit)?`,
    gradehope    = `What grade do you hope to get in DATA2002/2902?`,
    assess       = `When it comes to uni assessments, what do you tend to do?`,
    commute      = `On average, how long does it take you to get to campus? (mins) [1 direction]`,
    pay_rent     = `Do you pay rent?`,
    rent_amt     = `On average, how much rent do you pay per week? (dollars)`,
    work_hrs     = `On average, how many hours a week do you spend in paid employment? (hours)`,
    sleep_hrs    = `On average, how many hours sleep do you get per night? (hours)`,
    music        = `What is your favourite type of music?`,
    aliens       = `Do you believe in the existence of aliens? (Extraterrestial life)`,
    water        = `On average, how much water do you drink in a day? (litres)`,
    n_lang       = `How many spoken languages can you speak fluently?`,
    multitask    = `Do you multi-task in lectures? (ie do other tasks concurrently)`
  ) %>%
  filter(assess != "Other") %>%
  mutate(assess = factor(assess, levels = c("Immediately", "Gradually", "Cram")),
         multitask = factor(multitask, levels = c("no", "yes"), labels = c("No", "Yes")))

threshold <- 70

df_clean <- df %>% filter(work_hrs <= threshold)

ggplot(bind_rows(
    df %>% mutate(stage = "Before cleaning"),
    df_clean %>% mutate(stage = "After cleaning")
  ),
  aes(x = stage, y = work_hrs, fill = stage)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Before cleaning" = "#BFDCEB", "After cleaning" = "#F6D6A8")) +
  labs(
    title = "Weekly employment hours before and after cleaning",
    x = NULL, y = "Weekly employment hours", fill = NULL
  ) +
  theme_minimal()

write_csv(df_clean, "survey_cleaned.csv")
```

# Results

## RQ1: Work hours and cramming

**Research Question:** Is there a difference in weekly paid employment hours between students who cram before assessments and those who do not?

Weekly paid employment hours are often considered a key factor affecting student engagement. This research question examines whether work hours are associated with students' assessment strategy, specifically whether they tend to cram before deadlines. If time-pressured students are more likely to delay revision until the last minute, this insight could help the teaching team decide whether more flexible study support is needed for students with heavy work commitments. A Welch two-sample t-test is used to compare mean weekly work hours between Cram and Non-Cram students.

### Exploratory data analysis

```{r}
rq1_data <- df_clean %>%
  select(assess, work_hrs) %>%
  mutate(is_cram = if_else(assess == "Cram", "Cram", "Non-Cram"))
```

```{r}
rq1_data %>%
  group_by(is_cram) %>%
  summarise(
    n = n(),
    mean = mean(work_hrs),
    median = median(work_hrs),
    SD = sd(work_hrs),
    Q1 = quantile(work_hrs, 0.25),
    Q3 = quantile(work_hrs, 0.75),
    IQR = IQR(work_hrs),
    min = min(work_hrs),
    max = max(work_hrs)
  ) %>%
  knitr::kable(align = rep("c", 9))
```

```{r}
rq1_data %>%
  ggplot(aes(x = is_cram, y = work_hrs)) +
  geom_boxplot(
    aes(fill = is_cram),
    width = 0.5,
    outlier.shape = NA,
    alpha = 0.85
  ) +
  geom_jitter(
    aes(colour = "Individual observations"),
    position = position_jitter(width = 0.15, seed = 123),
    alpha = 0.35,
    size = 1.5
  ) +
  stat_summary(
    aes(colour = "Group mean"),
    fun = mean,
    geom = "point",
    shape = 18,
    size = 4
  ) +
  stat_summary(
    fun = mean,
    geom = "text",
    aes(label = after_stat(round(y, 2))),
    vjust = -0.75,
    size = 3.5
  ) +
  scale_fill_manual(
    name = "Cram status",
    values = c("Non-Cram" = "#BFE3DB", "Cram" = "#F3B5A7")
  ) +
  scale_colour_manual(
    name = NULL,
    values = c("Individual observations" = "#6F8FA3", "Group mean" = "#D9827A")
  ) +
  labs(
    x = NULL,
    y = "Weekly paid employment hours",
    title = "Weekly work hours by cramming behaviour",
    caption = "Diamonds and labels indicate group means."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "plain", hjust = 0.5),
    legend.position = "top",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )
```

Figure 1: Distribution of weekly paid employment hours by cramming status.

### Hypotheses

Let $\mu_1$ be the mean weekly work hours for Cram students and $\mu_2$ for Non-Cram students.

$H_0: \mu_1 = \mu_2$  
$H_1: \mu_1 \neq \mu_2$

### Assumptions

The Welch two-sample t-test assumes approximate normality within each group. We test this on the raw data using the Shapiro-Wilk test.

```{r}
cram_raw <- rq1_data %>% filter(is_cram == "Cram") %>% pull(work_hrs)
noncram_raw <- rq1_data %>% filter(is_cram == "Non-Cram") %>% pull(work_hrs)
shapiro_cram_raw <- shapiro.test(cram_raw)
shapiro_noncram_raw <- shapiro.test(noncram_raw)
 
tibble::tibble(
  Group = c("Cram", "Non-Cram"),
  Statistic = c(shapiro_cram_raw$statistic, shapiro_noncram_raw$statistic),
  p_value = c(shapiro_cram_raw$p.value, shapiro_noncram_raw$p.value)
) %>%
  mutate(
    Statistic = round(Statistic, 3),
    p_value = format.pval(p_value, digits = 3, eps = 0.001)
  ) %>%
  knitr::kable(align = c("l", "c", "c"), caption = "Shapiro-Wilk test on raw work hours")
```

Both groups have Shapiro-Wilk p-values far below 0.05, so the raw data does not satisfy the normality assumption. We apply a log(x+1) transformation and re-check normality below.

```{r}
cram_log <- log1p(cram_raw)
noncram_log <- log1p(noncram_raw)
shapiro_cram_log <- shapiro.test(cram_log)
shapiro_noncram_log <- shapiro.test(noncram_log)
 
tibble::tibble(
  Group = c("Cram", "Non-Cram"),
  Statistic = c(shapiro_cram_log$statistic, shapiro_noncram_log$statistic),
  p_value = c(shapiro_cram_log$p.value, shapiro_noncram_log$p.value)
) %>%
  mutate(
    Statistic = round(Statistic, 3),
    p_value = format.pval(p_value, digits = 3, eps = 0.001)
  ) %>%
  knitr::kable(align = c("l", "c", "c"), caption = "Shapiro-Wilk test on log-transformed work hours")

manual_skew <- function(x) {
  n <- length(x)
  m <- mean(x)
  (sum((x - m)^3) / n) / (sum((x - m)^2) / n)^(3/2)
}
```

Normality is still not fully satisfied after transformation. We compare skewness before and after the transformation to assess whether the distribution shape has nonetheless improved.

```{r}
skew_cram_raw <- manual_skew(cram_raw)
skew_noncram_raw <- manual_skew(noncram_raw)
skew_cram_log <- manual_skew(cram_log)
skew_noncram_log <- manual_skew(noncram_log)
tibble::tibble(
  Group = c("Cram", "Non-Cram"),
  `Raw skewness` = round(c(skew_cram_raw, skew_noncram_raw), 2),
  `Log skewness` = round(c(skew_cram_log, skew_noncram_log), 2)
) %>%
  knitr::kable(align = c("l", "c", "c"), caption = "Skewness before and after log transformation")
```

Skewness is substantially reduced after transformation for both groups. The residual non-normality is likely due to zero-inflation in the raw data, as shown by the clustering of points near zero in Figure 1, which no transformation can remove. Given the large sample size in both groups, the Central Limit Theorem still supports using a t-test.

```{r}
f_test <- var.test(cram_log, noncram_log)
 
tibble::tibble(
  Statistic = round(f_test$statistic, 3),
  p_value = format.pval(f_test$p.value, digits = 3, eps = 0.001)
) %>%
  knitr::kable(align = c("c", "c"), caption = "F test for equality of variances")
```

The F test shows no evidence of unequal variances. We proceed with a Welch two-sample t-test given the large sample size in both groups.

### Test statistic

```{r}
welch_test <- t.test(cram_log, noncram_log, var.equal = FALSE)
welch_test
```

The Welch two-sample t-test statistic is $t = 3.60$ (df = 167.28).

### P-value

The corresponding p-value is $p = 0.0004$. We calculate Cohen's d to quantify the size of this effect.

```{r}
#| label: rq1-effect-size
pooled_sd <- sqrt(((length(cram_log) - 1) * var(cram_log) +
                    (length(noncram_log) - 1) * var(noncram_log)) /
                    (length(cram_log) + length(noncram_log) - 2))
cohend <- (mean(cram_log) - mean(noncram_log)) / pooled_sd
cohend
```

Cohen's d = 0.372, indicating a small-to-medium effect size.

### Conclusion&Limitations
Since $p = 0.0004 < 0.05$, we reject $H_0$. Cram students report significantly higher weekly work hours (log scale) than Non-Cram students, with a small-to-medium effect (Cohen’s $d = 0.372$), suggesting that paid work is one of several factors associated with cramming. Although zero-inflation remained after log transformation, the large sample size supports the t-test approximation. As the data are observational, this association cannot establish that longer work hours cause cramming.

## RQ2: Multitasking and cramming

**Research Question:** Is multitasking during lectures associated with whether students cram before assessments?

Multitasking in lectures is usually read as disengagement. If it also tracks with cramming, the teaching team can treat lecture engagement and time management as one problem rather than two.

### Exploratory data analysis

```{r}
#| label: rq2-setup
rq2_data <- df %>%
  select(assess, multitask) %>%
  drop_na(assess, multitask) %>%
  mutate(
    is_cram = factor(
      if_else(assess == "Cram", "Cram", "Non-Cram"),
      levels = c("Non-Cram", "Cram")
    )
  )
```

```{r}
#| label: tbl-rq2-table
#| tbl-cap: "Cramming behaviour by multitasking status, with marginal totals."
cram_tab <- table(rq2_data$is_cram, rq2_data$multitask)

cram_tab %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  tibble::rownames_to_column("Preparation style") %>%
  knitr::kable(
    align = "c",
    col.names = c("Preparation style", "Does not multitask", "Multitasks", "Total")
  )
```

```{r}
#| label: rq2-computations
#| include: false
chi_test     <- chisq.test(cram_tab, correct = FALSE)
chi_statistic <- unname(chi_test$statistic)
chi_df        <- unname(chi_test$parameter)
chi_p         <- chi_test$p.value
min_expected  <- min(chi_test$expected)
n_total       <- sum(cram_tab)
cramers_v     <- sqrt(chi_statistic / n_total)

yes_counts   <- cram_tab[c("Cram", "Non-Cram"), "Yes"]
group_totals <- rowSums(cram_tab)[c("Cram", "Non-Cram")]
prop_test    <- prop.test(x = yes_counts, n = group_totals, correct = FALSE)

pct_cram    <- 100 * unname(prop_test$estimate[1])
pct_notcram <- 100 * unname(prop_test$estimate[2])
diff_ci     <- 100 * prop_test$conf.int
```

We use `df` rather than `survey_cleaned.csv`, whose `work_hrs` threshold also removes respondents this question does not need; rows are dropped only for missing `assess` or `multitask`, and `assess` is collapsed into cramming versus not. Multitasking was more common among students who cram (`r sprintf("%.1f", pct_cram)`%) than among those who do not (`r sprintf("%.1f", pct_notcram)`%), a difference of `r sprintf("%.1f", pct_cram - pct_notcram)` percentage points (@fig-rq2-eda). @fig-rq2-3level shows the two non-cramming categories behave similarly, supporting the collapse.

```{r}
#| label: fig-rq2-eda
#| fig-cap: "Proportion of students who multitask during lectures, by cramming behaviour."
rq2_data %>%
  count(is_cram, multitask) %>%
  group_by(is_cram) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = is_cram, y = prop, fill = multitask)) +
  geom_col(position = "fill", width = 0.5, alpha = 0.85) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(name = "Multitasks?",
                    values = c("No" = "#A8DADC", "Yes" = "#E76F51")) +
  labs(x = "Assessment preparation style", y = "Proportion of students",
       title = "Multitasking by cramming behaviour") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "plain", hjust = 0.5),
        legend.position = "top",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())
```

```{r}
#| label: fig-rq2-3level
#| fig-cap: "Proportion of students who multitask, by the original three-level preparation style."
rq2_data %>%
  count(assess, multitask) %>%
  group_by(assess) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = assess, y = prop, fill = multitask)) +
  geom_col(position = "fill", width = 0.5, alpha = 0.85) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(name = "Multitasks?",
                    values = c("No" = "#A8DADC", "Yes" = "#E76F51")) +
  labs(x = "Assessment preparation style (3 levels)",
       y = "Proportion of students",
       title = "Multitasking by assessment preparation style") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "plain", hjust = 0.5),
        legend.position = "top",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())
```

### Hypotheses

$H_0$: multitasking is independent of cramming among DATA2x02 students; $H_1$: the two are not independent.

### Assumptions

Responses from different students are treated as independent.

```{r}
#| label: tbl-rq2-expected
#| tbl-cap: "Expected cell counts under independence."
chi_test$expected %>%
  round(1) %>%
  as.data.frame.matrix() %>%
  tibble::rownames_to_column("Preparation style") %>%
  knitr::kable(
    align = "c",
    col.names = c("Preparation style", "Does not multitask", "Multitasks")
  )
```

The smallest expected count is `r sprintf("%.1f", min_expected)`, so all cells exceed 5 and the approximation is valid; no continuity correction is needed.

### Test statistic

```{r}
#| label: rq2-test
chi_test
```

Under $H_0$, $T=\sum_{i,j}(O_{ij}-E_{ij})^2/E_{ij}$ follows a $\chi^2$ distribution with `r chi_df` degree of freedom; the observed value is $t_0 = `r sprintf("%.2f", chi_statistic)`$.

### P-value

```{r}
#| label: tbl-rq2-effect-size
#| tbl-cap: "Effect size and confidence interval."
tibble::tibble(
  Measure = c("Cramér's V",
              "Difference in % multitasking (Cram - Non-Cram)"),
  Estimate = c(sprintf("%.3f", cramers_v),
               sprintf("%.1f pp", pct_cram - pct_notcram)),
  `95% CI` = c("-",
               sprintf("[%.1f, %.1f] pp", diff_ci[1], diff_ci[2]))
) %>%
  knitr::kable(align = c("l", "c", "c"))
```

The p-value is `r signif(chi_p, 3)`. Cramér's $V = `r sprintf("%.2f", cramers_v)`$ ($n = `r n_total`$); the 95% CI for the difference in multitasking rates is `r sprintf("%.1f", diff_ci[1])` to `r sprintf("%.1f", diff_ci[2])` percentage points.

### Conclusion

We reject $H_0$: multitasking and cramming are associated, though $V = `r sprintf("%.2f", cramers_v)`$ makes it one correlate of cramming rather than a defining feature.

### Limitations

Both variables are self-reported, so "cram" and "multi-task" may be applied inconsistently, and collapsing `assess` discards information. This observational, voluntary-response survey cannot establish causation — shared workload or procrastination could drive both — and may not generalise.

## RQ3: Paid employment and multitasking

**Research question:** Do students who multitask during lectures spend a
different average number of hours per week in paid employment compared with
students who do not multitask?

Paid employment represents an external time commitment for students. Students balancing more paid work with university study may exhibit different lecture behaviours. Having examined paid employment in relation to cramming and multitasking in relation to cramming, we investigate whether paid-employment hours also differ between students who do and do not multitask during lectures.

```{r}
#| label: rq3-setup
#| message: false
df_clean <- read_csv("survey_cleaned.csv") %>%
  mutate(
    work_hrs = as.numeric(work_hrs),
    multitask = factor(multitask, levels = c("No", "Yes"))
  ) %>%
  filter(is.na(work_hrs) | work_hrs <= 70)
```

```{r}
#| label: tbl-flipper
#| tbl-cap: "Summary statistics of weekly paid employment hours by multitasking status."

rq3_summary <- df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  group_by(multitask) %>%
  summarise(
    n = n(),
    mean = mean(work_hrs),
    median = median(work_hrs),
    SD = sd(work_hrs),
    Q1 = quantile(work_hrs, 0.25),
    Q3 = quantile(work_hrs, 0.75),
    IQR = IQR(work_hrs),
    min = min(work_hrs),
    max = max(work_hrs)
  )

knitr::kable(
  rq3_summary,
  align = rep("c", ncol(rq3_summary))
)
```
The boxplot in @fig-work-multitask compares weekly paid-employment hours between students who reported multitasking during lectures and those who did not. Students who reported multitasking had a higher mean weekly paid-employment time (8.02 hours) than those who did not multitask (6.23 hours), giving an observed mean difference of 1.79 hours per week. Both groups show visually right-skewed distributions, with many students reporting few or no hours of paid employment and a smaller number reporting substantially higher working hours. This observed difference in group means is examined formally using a permutation test.
```{r}
#| label: fig-work-multitask
#| fig-cap: "Distribution of weekly paid employment hours by multitasking status."

df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  ggplot(aes(x = multitask, y = work_hrs)) +

  geom_boxplot(
    aes(fill = multitask),
    width = 0.5,
    outlier.shape = NA,
    alpha = 0.4
  ) +

  geom_jitter(
    aes(colour = "Individual observations"),
    position = position_jitter(
      width = 0.15,
      seed = 123
    ),
    alpha = 0.35,
    size = 1.5
  ) +

  stat_summary(
    aes(colour = "Group mean"),
    fun = mean,
    geom = "point",
    shape = 18,
    size = 4
  ) +
  
  stat_summary(
    fun = mean,
    geom = "text",
    aes(label = after_stat(round(y, 2))),
    vjust = -0.75,
    size = 3.5
  ) +

  scale_fill_manual(
    name = "Multitasking",
    values = c(
      "No" = "#A8DADC",
      "Yes" = "#F4A261"
    )
  ) +
  
  scale_colour_manual(
    name = NULL,
    values = c(
      "Individual observations" = "#457B9D",
      "Group mean" = "red"
    )
  ) +

  labs(
    x = "Multitasking during lectures",
    y = "Weekly paid employment (hours)",
    title = "Weekly paid employment hours by multitasking status",
    caption = "Diamonds and labels indicate group means."
  ) +

  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )
```

```{r}
#| label: rq3-permutation-test
#| include: false

rq3_data <- df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  mutate(multitask = factor(multitask, levels = c("No", "Yes")))

welch_stat <- function(hours, group) {
  hours_yes <- hours[group == "Yes"]
  hours_no  <- hours[group == "No"]

  (mean(hours_yes) - mean(hours_no)) /
    sqrt(var(hours_yes) / length(hours_yes) +
           var(hours_no) / length(hours_no))
}

mean_yes <- mean(rq3_data$work_hrs[rq3_data$multitask == "Yes"])
mean_no  <- mean(rq3_data$work_hrs[rq3_data$multitask == "No"])
mean_difference <- mean_yes - mean_no
t_observed <- welch_stat(rq3_data$work_hrs, rq3_data$multitask)

set.seed(2002)
B <- 10000

t_permuted <- replicate(
  B,
  welch_stat(
    rq3_data$work_hrs,
    sample(rq3_data$multitask, replace = FALSE)
  )
)

n_yes <- sum(rq3_data$multitask == "Yes")
n_no  <- sum(rq3_data$multitask == "No")
sd_yes <- sd(rq3_data$work_hrs[rq3_data$multitask == "Yes"])
sd_no  <- sd(rq3_data$work_hrs[rq3_data$multitask == "No"])
pooled_sd <- sqrt(
  ((n_yes - 1) * sd_yes^2 + (n_no - 1) * sd_no^2) /
    (n_yes + n_no - 2)
)
cohens_d <- mean_difference / pooled_sd
```

```{r}
#| label: fig-rq3-permutation
#| fig-cap: "Permutation null distribution of the Welch t-statistic. Dashed red lines indicate the observed cut-offs at plus and minus the absolute observed statistic."

tibble(t_permuted = t_permuted) %>%
  ggplot(aes(x = t_permuted)) +
  geom_histogram(
    bins = 45,
    fill = "#8ECAE6",
    colour = "white"
  ) +
  geom_vline(
    xintercept = c(-abs(t_observed), abs(t_observed)),
    colour = "#D62828",
    linewidth = 0.9,
    linetype = "dashed"
  ) +
  labs(
    x = "Permuted Welch t-statistic",
    y = "Number of permutations",
    title = "Permutation distribution under the null hypothesis"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```

The two-sided permutation p-value is calculated by counting simulated
statistics at least as extreme as the observed statistic in either direction:

```{r}
#| label: rq3-p-value

extreme_count <- sum(abs(t_permuted) >= abs(t_observed))

# Plus-one correction for a Monte Carlo permutation p-value
p_permutation <- (extreme_count + 1) / (B + 1)

tibble(
  extreme_permutations = extreme_count,
  total_permutations = B,
  permutation_p_value = p_permutation
) %>%
  knitr::kable(
    format = "html",
    digits = c(0, 0, 4),
    col.names = c(
      "Extreme permutations",
      "Total permutations",
      "Permutation p-value"
    ),
    align = c("c", "c", "c"),
    table.attr = 'class="table table-bordered table-striped" style="width: auto; margin-left: auto; margin-right: auto;"'
  )
```

We formally test the difference using a two-sided permutation test with the
Welch $t$-statistic as the resampling statistic.

1. **Hypotheses:** Let $\mu_Y$ and $\mu_N$ be the population mean weekly
   paid-employment hours for students who do and do not multitask,
   respectively. $H_0:\mu_Y-\mu_N=0$ versus
   $H_A:\mu_Y-\mu_N\neq0$.

2. **Assumptions:** Responses from different students are treated as
   independent. Under $H_0$, multitasking labels are assumed exchangeable. A
   normal population is not required for the permutation p-value, although the
   mean-based statistic remains sensitive to large values.

3. **Test statistic:**
   $T=(\bar X_Y-\bar X_N)/\sqrt{S_Y^2/n_Y+S_N^2/n_N}$. The observed difference
   is `r sprintf("%.2f", mean_difference)` hours per week and
   $t_0=`r sprintf("%.3f", t_observed)`$.

4. **Permutation procedure:** With `set.seed(2002)`, the Yes/No labels were
   randomly reassigned $B=10{,}000$ times while employment hours were held
   fixed. The Welch statistic was recalculated each time to form the null
   distribution in @fig-rq3-permutation.

5. **P-value:** For the two-sided test, permutations satisfying
   $|T_b^*|\geq|t_0|$ were counted. The plus-one estimate is
   $p_{\mathrm{perm}}=`r sprintf("%.4f", p_permutation)`$.

6. **Decision:** Since $p<0.05$, we reject $H_0$. There is evidence that mean
   paid-employment hours differ by multitasking status; multitasking students
   worked `r sprintf("%.2f", mean_difference)` more hours per week on average
   in this sample. However, Cohen's $d=`r sprintf("%.2f", cohens_d)`$ indicates
   a small effect, so its practical importance is limited. Both variables are
   self-reported, the 70-hour cleaning threshold is analyst-chosen, and
   confounding may remain. As this is an observational survey of DATA2x02
   students, the result cannot establish causation or automatically generalise
   to other students.

# Conclusion

[Integrate the three findings into one story]

[Implications for client]

[Overall limitations / no causal inference]


# AI Usage Statement

[...]


# ISI Professional Ethics Statement

[...]


# References
